\section{Definition of an Event Tree}
\label{sec:event_tree_definition}

An \acrfull{et} unravels how a single \emph{initiating event} (\(I\)) can branch into multiple possible \emph{end-states} (\(X\)) through a sequence of \emph{functional} (or conditional) events. Each branch captures the success or failure of an important \emph{functional event} (e.g.\ a safety barrier or operator intervention). By following all possible paths, one can systematically account for each final outcome \(X_j\). Figure~\ref{fig:event_tree_example} provides a schematic view of this process for an initiating event \(I\) and two subsequent functional events, \(F_1\) and \(F_2\). Each terminal node (leaf) corresponds to a distinct end-state, denoted \(X_1, X_2, \ldots, X_n\). Though this illustration is intentionally simple, more complex systems may include numerous functional events, each branching into further outcomes.

\input{2_foundations/definition/figs/et_example}


At the highest conceptual level, an event tree is a collection of conditional outcomes. Let \(n\) be a positive integer, and let \(j\) range over some index set of end-states~\(J\). Then the scenario pathways can be defined as:
\begin{equation}
\label{eq:event_tree_gamma}
    \Gamma
    \;=\;
    \Bigl\{
        \langle
            I,\,
            F_1,\,
            F_2,\,
            \dots,\,
            F_n,\,
            X_j
        \rangle
        :\,
        j \in J
    \Bigr\},
\end{equation}
where:
\begin{itemize}
    \item \(I\) is the \textbf{initiating event}. In a nuclear system, this could be an abnormal occurrence such as a coolant pump trip or an unplanned reactivity insertion.
    \item \(F_k\) (\(k=1,\ldots,n\)) denotes the \(k\)th \textbf{functional (conditional) event}, which may succeed (\(F_k^{\text{succ}}\)) or fail (\(F_k^{\text{fail}}\)). Typically, each \(F_k\) depends on the outcomes \(F_1,\ldots,F_{k-1}\).
    \item \(X_j\) is an \textbf{end-state}, describing the final outcome along a particular branch. End-states might indicate safe shutdown, core damage, or a radiological release.
\end{itemize}
Each tuple \(\langle I, F_1, \dots, F_n, X_j\rangle\) in \(\Gamma\) encapsulates a distinct scenario pathway. In the broader context of the risk triplet, such a pathway corresponds to \(S_i\), the possibility of something going wrong, while the associated probability and consequences map directly to \(L_i\) and \(X_i\).