\subsection{Decision Diagrams}
\label{sec:decision_diagrams}

Decision diagrams provide a powerful, directed-graph-based representation of logical or Boolean functions. Their roots can be traced back to branching program ideas explored by Lee and Akers in the 1950s--1970s, but major refinement and widespread adoption occurred after Bryant's seminal work on \emph{Ordered Binary Decision Diagrams (OBDDs)} in 1986. In reliability analysis, formal verification, and combinational circuit design, decision diagrams frequently offer more computationally tractable methods than naïve enumeration of all input patterns.

This section introduces the basic concepts of \emph{Binary Decision Diagrams (BDDs)} and \emph{Zero-Suppressed Decision Diagrams (ZDDs)}, along with the special class of \emph{Ordered} and \emph{Reduced} BDDs that guarantee a canonical (unique) form under fixed conditions. We emphasize:
\begin{itemize}
\item The structure and interpretation of BDDs as directed acyclic graphs (DAGs).
\item The notion of an \emph{ordered} BDD, imposing a strict arrangement on variable testing.
\item Techniques for \emph{reducing} BDDs into smaller yet equivalent graphs by merging or removing redundant parts.
\item The main principles of Zero-Suppressed Decision Diagrams, designed for efficiently encoding sparse sets or combinatorial families.
\end{itemize}

Earlier, we saw that \emph{event trees} and \emph{fault trees} can be merged into a single directed acyclic graph to represent complex system dependencies. BDDs and ZDDs, by contrast, focus more narrowly on Boolean functions, providing specialized node-splitting and merging operations to systematically capture logical behavior. Despite differing motivations, both families of DAG-based representations benefit from the avoidance of cycles and the ability to encode large models in a structured form.


\input{2_foundations/knowledge_compilation/transformation/as_dd/bdd}
\input{2_foundations/knowledge_compilation/transformation/as_dd/zdd}
\input{2_foundations/knowledge_compilation/transformation/as_dd/sdd}
