\chapter{Introduction}

\section{\color{blue}{Project Motivation}}
The development and use of PRAs in the nuclear industry have revolutionized our approach to design, reliability, and safety. Since the release of WASH-1400 in 1975 and NUREG-1150 in 1990, the PRA applications have diversified and become more computationally demanding, nevertheless the tools that we use to perform such assessments have not kept up with the technological advancements in high-performance computing and computer science applications. Now, although there is a need for supporting real-time decisions at the nuclear plants using PRAs, this is not practical using the computationally strained legacy PRA tools currently available since they were developed and are maintained using technologies that are nowadays deprecated. The legacy PRA tools, although still perform well on internal events PRA models that have reasonable sizes, suffer greatly, both in terms of speed and memory requirements, when challenged by more sophisticated models such as single hazard PRAs, and especially multi-hazard PRAs. Therefore, a major redesign of the PRA tools is necessary starting from the computational engine capabilities, backend services to handle large PRA models, and a user-friendly PRA frontend that can automatically generate results and documentation necessary to inform non-PRA experts.

\section{\color{blue}{Project Scope}}
The scope of this project is to design, implement, and evaluate an open source, parallel, and distributed web-based PRA platform that addresses the limitations of legacy PRA tools. The platform is intended to support a wide range of PRA applications, from traditional internal events models to advanced multi-hazard and multi-unit scenarios. The project encompasses the development of a modular architecture that integrates a web-based client for model editing and visualization, RESTful backend APIs for data management and orchestration, and a distributed job queue system for high-throughput risk quantification. The platform is designed to be extensible, supporting integration with multiple quantification engines and enabling interoperability with industry-standard tools through the adoption of the OpenPRA schema and compatibility with the OpenPSA Model Exchange Format (MEF). The project also includes the development of automated benchmarking and model translation workflows, as well as the implementation of features to facilitate collaborative model development, transparent risk communication, and reproducible analysis. The ultimate goal is to provide a robust, scalable, and user-friendly solution that meets the evolving needs of PRA practitioners and stakeholders in the nuclear industry.

\section{\color{blue}{Project Objectives}}
The main objective of this work is to develop, demonstrate, and evaluate a probabilistic risk assessment (PRA) software platform needed to address the major challenges of the current legacy PRA tools, such as better quantification speed, integration of multi-hazard models into traditional PRAs, and model modification simplification and documentation automation. To achieve the main objective, we will first perform benchmarking and profiling of current PRA tools, such as SCRAM [1] and US NRC's SAPHIRE [2, Vol. 1] , to investigate the current bottlenecks in the quantification speed and memory requirements. Secondly, we will design, implement, and benchmark a PRA software platform based on a web-based stack using the latest technologies available to overcome the mentioned challenges. Finally, we will evaluate the performance gains of this framework by modeling and quantifying large PRA models that would have been too expensive to run using the legacy PRA tools.

\section{\color{blue}{Objectives of this Report}}
This report documents the design, implementation, and evaluation of the open source, parallel, and distributed web-based PRA platform developed under this project. The objectives of this report are to:
\begin{itemize}
    \item Present the motivation, scope, and objectives of the project in the context of current challenges in PRA technology.
    \item Describe the architectural design and technical implementation of the platform, including its modular components and integration strategies.
    \item Summarize the benchmarking and evaluation of the platform against existing PRA tools, highlighting improvements in computational efficiency, scalability, and model interoperability.
    \item Identify current limitations and outline a roadmap for future research and development to address remaining gaps and advance the platform toward industry readiness.
    \item Provide guidance and best practices for PRA practitioners and stakeholders interested in adopting or contributing to the platform.
\end{itemize}

\input{1_introduction/outcomes/_}

\input{1_introduction/howto_recreate}

\input{1_introduction/layout}