\begin{abstract}
This report presents the development and current status of an open source, parallel, and distributed web-based probabilistic risk assessment (PRA) platform designed to support real time, risk-informed operational decisions in nuclear power plants. The platform implements a modular architecture comprising a web-based client, RESTful backend APIs, distributed job queues, and a scalable worker pool capable of leveraging both CPU and GPU resources. Model representation is standardized using an extensible OpenPRA schema, supporting both single and multi-hazard PRA models, and enabling interoperability with industry standard tools such as SAPHIRE. Benchmarking against established PRA engines demonstrates significant improvements in quantification speed and scalability, particularly for large and complex models. The platform supports benchmarking, model translation, and integration with multiple quantification engines, and provides a foundation for collaborative model development and transparent risk communication. However, several advanced features remain under development, including full version control, real time collaborative editing, adaptive job scheduling, and comprehensive support for all OpenPSA Model Exchange Format (MEF) elements. Ongoing work focuses on extending solver support, improving rare event quantification, and enhancing data management for very large models. The platform’s open architecture and extensibility position it as a strong foundation for future research and industry adoption, with the goal of enabling transparent, and reproducible risk quantification for real time operational decision support in the nuclear sector.
\end{abstract}