\begin{acknowledgments}
This research was supported by the Department of Energy's (DOE) Nuclear Energy University Program (NEUP) under project number 21-24004. The authors gratefully acknowledge the contributions of several individuals whose expertise and efforts were instrumental to the success of this work.

We thank Akram Batikh (NCSU NE PRA Group) for his development of the multi-hazard and aftershocks probabilistic risk assessment (PRA) model using OpenMHA, a subset of the OpenPRA platform. We also appreciate the assistance of Alp Tezbasaran (NCSU NE PRA Group) for his support in troubleshooting benchmarking scripts, which improved the reliability and reproducibility of our computational experiments.

We further thank Rahul Agarwal (TrueClean Energy Technologies) for his implementation of the OpenPRA schema and the development of TypeScript tooling for documentation and validation. His work on the shared-types package schema, available in the project monorepo, was essential for ensuring model interoperability and validation.

The authors are grateful for the collaborative environment and technical insights provided by all contributors.

\color{red}{Thank the MHTGR modelers Priyanka Pandit, Mostafa Hamza, Dave Blanchard, Rawan Mustafa.}

\color{red}{Thank the G-SCFR modelers Asmaa Farag (also author), Amy Whitley and Elijah Rushing.}

\color{red}{Thank Yahya Alzarahni for discussions about event sequence modeling that helped bridge the gap between mutually exclusive event sequences and the disjuntive normal form.}


\end{acknowledgments}